Identification and analysis of pleiotropic expression quantitative trait loci

### 2023-05-08 doc/presentation_230511_centuriday2023/presentation_centuri2023_marseille.tex
DOC centuri day

        \begin{itemize}
            \item Phenotypic differences between human populations are mostly coded in frequent genetic variants
            \item These genetic variants also cause different disease susceptibility
            \item Large datasets of associations have shown that many genetic variants change simultaneously susceptibility to many diseases, ie. the variants are pleiotropic
        \end{itemize}
%
        \vfill
%
        What is the molecular mechanism and effect of tissue of these variants?

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Expression quantitative trait loci (eQTL)}

        \begin{itemize}
            \item Quantitative trait loci (QTL) are variants associated with changes of quantitative phenotypes
            \item Expression (eQTL) are variants associated with changes of gene expression in a given tissue
            \item eQTLs provide a putative molecular mechanism of a variant
        \end{itemize}
        \vfill
        What are the properties of pleiotropic eQTLs?

@@

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Summary}

>>>>

        \begin{itemize}
            \item Differences between humans coded in 80 x 10$^6$ frequent genetic variants
            \item Genetic variants change susceptibilities for thousands of common traits and diseases
            \item For example, breast cancer:
        \end{itemize}
        %
        \begin{figure}
            \begin{center}
                \includegraphics[width=0.5\textwidth]{\presentationgoldparispath/41586_2017_Article_BFnature24284_Fig1_HTML.jpg}
            \end{center}
            \caption{Genetic variant associations with breast cancer risk (DOI:10.1038/nature24284).}
        \end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Effect of genetic variants on gene expression: eQTLs}

        \begin{itemize}
            \item Quantitative trait loci (QTL) variants associated with changes of quantitative phenotypes
            \item Expression (eQTL) are variants associated with changes of gene expression in a given tissue
            \item eQTLs provide a putative molecular mechanism of a variant
        \end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Pleiotropic genetic variants}

        \begin{itemize}
            \item Large datasets of associations have shown that many genetic variants change simultaneously susceptibility to many diseases, ie. the variants are pleiotropic
        \end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Summary of previous and my work here}

        \begin{center}
            \includegraphics[width=0.6\textwidth]{fig/graphical_intro_approach.drawio.png}
        \end{center}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Most pleiotroic eQTLs}

        \begin{itemize}
            \item Example pleiotropic eQTLs in different cytobands
            \item The representative gene is the eQTL gene with the highest PubMed publication count
        \end{itemize}

% full size table is table
        \begin{table}[!tbp]
    \centering
    \scriptsize

    \csvreader[separator=tab,
    tabular=crcp{0.4\textwidth},
    late after last line=\\\hline,
    head,
    table head=\\\hline \bfseries Chrom. & \bfseries Pos (hg38) & \bfseries Gene marker& \bfseries Trait categories\\\hline,
    ]{\mspath/cmpt_count_per_rsid.py/count_per_rsid_gwas_ms.tsv}{}% use head of csv as column names
        {\csvcoli\ & \csvcolii\ & \csvcolv\ & \csvcolvi}% specify your coloumns here
            \vspace{15pt}\label{tab:pleitropic_variants}
        \end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{frame}
        \frametitle{Summary of my work}
